---
title: Academic Self-Concept Project
author: 'Silas Jackson'
date: '2021-04-24'
summary: A survey of students' academic self-concept, college major, and careerism.
slug: Academic-Self-Project
linktitle: "Academic Self and Careerism"
output:
  blogdown::html_page:
    
    toc: true
menu:
  docs:
    parent: "Projects"
    weight: 1
type: docs
editor_options: 
  chunk_output_type: inline
lastmod: '2021-04-24T23:09:32-04:00'
featured: no
projects: []
---
# **Overview**

## **Background**

  One part of James Daniel Lee’s argument in Which Kids Can "Become" Scientists? Effects of Gender, Self-Concepts, and Perceptions of Scientists is that girls are less likely to pursue science careers because they feel their self-concept(s) are not in line with the expectation of what a scientist is like. This suggests that, at least in part, students tend to pursue careers which they feel are congruent with their self-concept. However, this view does not account for whether college students consider their education a means to pursue their passions and express a part of their “true” self, or as a path to obtaining a lucrative career. Further, research on academic self-concept almost exclusively focuses on achievement as its main indicator or correlate. 


  I argue that achievement is only one aspect of academic self-concept and research on this topic must account for academic gratification. That said, I seek to (1) examine whether students college major is more influenced by careerist goals or self-concepts, (2) explore how others’ perceptions (reflected appraisals) affect their self-concept and overall satisfaction with their major, (3) reveal if students who choose their major based on self-concept are more or less satisfied with their major, and (4) investigate students’ attitudes about their perceptions of their future success as it relates to the conclusions above.

## **Research Questions**

* What is the role of self-concept in students’ choice of college major? 
* What is the role of reflected appraisals in satisfaction with college major?
* Are students more likely to choose a major based on self-concept or career outcomes?
* Are students who choose a major based on self-concept more or less satisfied than those who choose a major on career outcomes?
* What is the bridge between academic self-concept and general self-concept?


## **Methods**

### Survey
Download the survey [here.](/media/Survey on Academic Self.pdf)



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, cache = TRUE)


load("AcadSelfData.Rdata")
load("AcadLikertData.Rdata")


AcadSelfData$GPA <- plyr::mapvalues(AcadSelfData$GPA, from = "I don't have a GPA (First-semester freshmen)", to = "I don't know.") 
AcadSelfData$GPA <- plyr::mapvalues(AcadSelfData$GPA, from = "4", to = "4.0")

AcadLikertData2 <- AcadLikertData[, c(1:6, 8:18)]
cols <- colnames(AcadLikertData2)


```

```{r libraries, include = FALSE}
library(tidyverse)
library(likert)
library(ggthemes)
library(extrafont)
library(plotly)
library(viridis)
library(ggpubr)
library(data.table)
library(rcompanion)
library(DescTools)
library(summarytools)
library(ordinal)
library(MASS)
library(qqplotr)
library(widgetframe)
library(htmltools)
library(htmlwidgets)
library(htmlTable)
```




```{r Theme setup, include = FALSE}

theme_update(text = element_text(family = "Rockwell"),
        plot.title = element_text(size = 15, face = "bold"),
        plot.title.position = "plot",
        panel.background = element_rect(fill = "white"),
        panel.grid.minor.y = element_line(linetype = "dotted", color = "grey50"),
        panel.grid.major.y = element_line(color = "grey90"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor = element_blank(),
        plot.margin = unit(c(0.5, 0.5, 0.5, 0.5), units = "cm"))



threeColors <- c("#5b0909", "#34457c", "#0fb040")

fourColors <- c("#5b0909", "#571b43", "#34457c", "#0fb040")

fiveColors <- c("#5b0909", "#571b43", "#34457c", "#008a9e", "#0fb040")

sevenColors <- c("#5b0909", "#5f0720","#571b43", "#34457c", 
                 "#00558d","#008a9e","#0fb040")

eightColors <- c("#5b0909", "#5f0720","#571b43", "#4c2651", 
                 "#34457c", "#00558d","#008a9e","#0fb040")

nineColors <- c("#5b0909", "#5e0f33", "#571b43", "#4c2651", 
                  "#453568", "#34457c", "#0071a4", "#008a9e", 
                   "#0fb040")

twelveColors <- c("#5b0909", "#5f0720", "#5e0f33", "#571b43", "#4c2651", 
                  "#453568", "#34457c", "#00558d", "#0071a4", "#008a9e", 
                  "#00a07a", "#0fb040")

```




# **Sample Characteristics**

```{r Sample Characteristics, message=FALSE, warning=FALSE}
ggplot(AcadSelfData, aes(Gender, fill = Gender)) +
  geom_bar(show.legend = FALSE) +
  geom_text(aes(label = ..count..), stat = "count", vjust = -1)+
  labs(title = "Sample Characteristics: Gender",
       y = NULL) +
  scale_fill_manual(values = threeColors) +
  ylim(0, 550)


ggplot(AcadSelfData, aes(ClassLvl, fill = ClassLvl)) +
  geom_bar(show.legend = FALSE) +
  geom_text(aes(label = ..count..), stat = "count", vjust = -1) +
  labs(title = "Sample Characteristics: Class Level",
       y = NULL,
       x = "Class Level") +
  scale_fill_manual(values = fourColors) +
  ylim(0, 225)


ggplot(AcadSelfData, aes(Majorgrouped2, fill = Majorgrouped2)) +
  geom_bar(show.legend = FALSE) +
  geom_text(aes(label = ..count..), stat = "count", hjust = -0.5) +
  labs(y = NULL,
       x = NULL,
       title = "Sample Characteristics: Area of Study",
       caption = "'Pre-prof.' indicates pre-professional fields such as pre-dental and pre-medical.") +
  scale_fill_manual(values = sevenColors) +
  ylim(0, 150) +
  coord_flip()


```



```{r}
k <- ggplot(AcadSelfData, aes(STEMmaj, fill = Majorgrouped2, text = paste("STEM Major?:", STEMmaj,
                                                                          "<br>Area of Study:", Majorgrouped2))) +
  geom_bar(position = "dodge") +
  labs(title = "Do you consider your major a STEM major?",
       y = NULL,
       x = NULL,
       fill = NULL) +
  scale_fill_viridis(discrete = TRUE) +
  theme(legend.position = "bottom",
        legend.direction = "vertical",
        legend.key.width = unit(3, "cm"),
        legend.key.height = unit(3, "mm")) +
  guides(fill = guide_legend(nrow = 1,
                      label.position = "right"))


ggplotly(k,
         tooltip = c("text", "count"))
```




```{r}
ggplot(AcadSelfData, aes(GPA, fill = GPA)) +
  geom_bar(show.legend = FALSE) +
  geom_text(aes(label = ..count..), stat = "count", vjust = - 0.5) +
  labs(title = "Sample Characteristics: GPA",
       y = NULL) +
  ylim(0, 200) +
  scale_fill_manual(values = eightColors)



ggplot(AcadSelfData, aes(GPA2, fill = GPA2)) +
  geom_bar(show.legend = FALSE) +
  geom_text(aes(label = ..count..), stat = "count", vjust = - 0.5) +
  labs(title = "Sample Characteristics: GPA",
       y = NULL,
       x = "GPA") +
  scale_fill_manual(values = threeColors) +
  ylim(0, 425)

```




```{r Likert Scale Statements, fig.height=8, fig.width=8}
p <- plot(likert(AcadLikertData), col = sevenColors) +
  labs(title = "Level of Agreement Questions") +
  theme(legend.position = "none",
        legend.direction = "horizontal",
        plot.title.position = "plot",
        legend.title = element_blank(),
        legend.justification = "left",
        legend.key.width = unit(0.1, "mm"),
        legend.key.height = unit(0.35, "cm"),
        legend.key.size = unit(0.1, "mm"),
        legend.text = element_text(size = 8),
        plot.title = element_text(size = 15, hjust = 0.1),
        legend.background = element_blank(),
        panel.grid.major.y = element_line(linetype = "dotted")) +
  guides(fill = guide_legend(label.position = "bottom",
                             nrow = 1))

ggplotly(p)

```

```{r}

round(PairApply(AcadLikertData2, CramerV, symmetric = FALSE), 4) %>% 
  data.table()

```




# **Analysis**



```{r}
ggplot(AcadSelfData, aes(CareerImportant, fill = CareerImportant)) +
  geom_bar(show.legend = FALSE) +
  geom_text(aes(label = ..count..), stat = "count", vjust = -0.5) +
  labs(title = "How important was having a high-paying career\nin choosing your current major?",
       x = NULL,
       y = NULL) +
  ylim(0, 325) +
  scale_fill_viridis(discrete = TRUE,
                     direction = -1)

ggplot(AcadSelfData, aes(PassionImportant, fill = PassionImportant)) +
  geom_bar(show.legend = FALSE) +
  geom_text(aes(label = ..count..), stat = "count", vjust = -0.5) +
  labs(title = "How important was having a career you are passionate\nabout in choosing your current major?",
       x = NULL,
       y = NULL) +
  ylim(0, 450) +
  scale_fill_viridis(discrete = TRUE,
                     direction = -1)




```




```{r}

more_imp <- AcadSelfData[,c(16:17) ]


for(i in 1:ncol(more_imp)){
  
  levels(more_imp[, i]) <- c("1", "2", "3", "4", "5")
  
}


more_imp <- as.data.frame(apply(more_imp, 2,as.numeric))



more_imp <- more_imp %>% 
  mutate(impscore = case_when(PassionImportant > CareerImportant ~ "Passion",
                              CareerImportant > PassionImportant ~ "Career",
                              CareerImportant == PassionImportant ~ "Equal"))

more_imp <- more_imp %>% 
  mutate(v = PassionImportant - CareerImportant)


AcadSelfData$more_imp <- more_imp$impscore
AcadSelfData$impscore <- more_imp$v

AcadSelfData <- AcadSelfData %>% 
  mutate(ImSTEM = case_when(STEMmaj == "Yes" ~ "Yes",
                            STEMmaj == "Maybe/Unsure" ~ "Maybe/Unsure",
                            STEMmaj == "No" ~ "No",
                            TRUE ~ "No"))

```


```{r}
ggline(AcadSelfData, x = "Majorgrouped2", y = "impscore", 
       add = c("mean_se", "jitter"),
       ylab = "Score",
       xlab = "",
       title = "",
       point.color = "red",
       point.size = 1,
       color = "black",
       add.params = list(alpha = 0.5)) +
       rotate_x_text(angle = 45)



```


```{r echo=FALSE}
table1 <- AcadSelfData %>%
  group_by(Majorgrouped2) %>%
  rename("Major?" = "Majorgrouped2") %>% 
  summarise(n = n(), 
            mean = round(mean(impscore, na.rm = TRUE), 2), 
            sd = round(sd(impscore, na.rm = TRUE), 2),
            stderr = round(sd/sqrt(n), 2),
            LCL = round((mean - qt(1 - (0.05 / 2), n - 1) * stderr), 2),
            UCL = round((mean + qt(1 - (0.05 / 2), n - 1) * stderr), 2),
            median = median(impscore, na.rm = TRUE),
            min = min(impscore, na.rm = TRUE), 
            max = max(impscore, na.rm = TRUE),
            IQR = IQR(impscore, na.rm = TRUE),
            W.Stat = round((shapiro.test(impscore)$statistic), 4),
            p.value = round((shapiro.test(impscore)$p.value), 2)) %>% 
      data.table::data.table()


table1



broom::tidy(kruskal.test(impscore ~ Majorgrouped2, AcadSelfData)) %>%
  data.table::data.table()

```


```{r}


ggline(AcadSelfData, x = "ImSTEM", y = "impscore", 
       add = c("mean_se", "jitter"),
       ylab = "Score",
       xlab = "",
       title = "Major",
       point.color = "red",
       point.size = 1,
       color = "black",
       add.params = list(alpha = 0.5))




AcadSelfData %>%
  group_by(STEMmaj) %>%
  rename("STEM?" = "STEMmaj") %>% 
  summarise(n = n(), 
            mean = round(mean(impscore, na.rm = TRUE), 2), 
            sd = round(sd(impscore, na.rm = TRUE), 2),
            stderr = round(sd/sqrt(n), 2),
            LCL = round((mean - qt(1 - (0.05 / 2), n - 1) * stderr), 2),
            UCL = round((mean + qt(1 - (0.05 / 2), n - 1) * stderr), 2),
            median = median(impscore, na.rm = TRUE),
            min = min(impscore, na.rm = TRUE), 
            max = max(impscore, na.rm = TRUE),
            IQR = IQR(impscore, na.rm = TRUE),
            W.Stat = round((shapiro.test(impscore)$statistic), 4),
            p.value = round((shapiro.test(impscore)$p.value), 2)) %>% 
  data.table::data.table()



broom::tidy(kruskal.test(impscore ~ ImSTEM, AcadSelfData)) %>% 
  data.table() 

```




```{r}
ggplot(AcadSelfData, aes(passion_vs_career, fill = passion_vs_career)) +
  geom_bar(show.legend = FALSE) +
  geom_text(aes(label = ..count..), stat = "count", hjust = -0.5) +
  labs(title = "What is most important, going to ollege to\nget a good career or to explore your passion?",
       x = NULL,
       y = NULL) +
  coord_flip() +
  ylim(0, 300) +
  theme(plot.title.position = "plot")
```






```{r}
ggplot(AcadSelfData, aes(true_self, fill = true_self)) +
  geom_bar(show.legend = FALSE) +
  geom_text(aes(label = ..count..), stat = "count", vjust = -0.5) +
  labs(title = "My major is a reflection of my true self.",
       x = NULL,
       y = NULL) +
  scale_fill_viridis(discrete = TRUE,
                     option = "E") +
  ylim(0, 275)



ggplot(AcadSelfData, aes(fulfilled, fill = fulfilled)) +
  geom_bar(show.legend = FALSE) +
  geom_text(aes(label = ..count..), stat = "count", vjust = -0.5) +
  labs(title = "How fulfilled do you feel in your major?",
       x = NULL,
       y = NULL) +
  scale_fill_viridis(discrete = TRUE,
                     option = "E") +
  ylim(0, 300)



ggplot(AcadSelfData, aes(fit_in, fill = fit_in)) +
  geom_bar(show.legend = FALSE) +
  geom_text(aes(label = ..count..), stat = "count", vjust = -0.5) +
  labs(title = "I feel like I fit in easily with other students in my major.",
       x = NULL,
       y = NULL) +
  scale_fill_viridis(discrete = TRUE,
                     option = "E") +
  ylim(0, 225)



ggplot(AcadSelfData, aes(overall_satisfied, fill = overall_satisfied)) +
  geom_bar(show.legend = FALSE) +
  geom_text(aes(label = ..count..), stat = "count", vjust = -0.5) +
  labs(title = "Overall, How satisfied or unsatisfied are you\nwith your college academic experience?",
       x = NULL,
       y = NULL) +
  scale_fill_viridis(discrete = TRUE,
                     option = "E") +
  ylim(0, 350)




ggplot(AcadSelfData, aes(rewarding, fill = rewarding)) +
  geom_bar(show.legend = FALSE) +
  geom_text(aes(label = ..count..), stat = "count", vjust = -0.5) +
  labs(title = "Being a student is a rewarding experience.",
       x = NULL,
       y = NULL) +
  scale_fill_viridis(discrete = TRUE,
                     option = "E") +
  ylim(0, 225)




ggplot(AcadSelfData, aes(money_issue, fill = money_issue)) +
  geom_bar(show.legend = FALSE) +
  geom_text(aes(label = ..count..), stat = "count", vjust = -0.5) +
  labs(title = "If money were not an issue, I would have chosen a different major.",
       x = NULL,
       y = NULL) +
  scale_fill_viridis(discrete = TRUE,
                     option = "E") +
  ylim(0, 275)


```

## **Limitations**


# **References**



