---
title: Academic Self-Concept Project
author: 'Silas Jackson'
date: '2021-04-24'
summary: A survey of students' academic self-concept, college major, and careerism.
slug: Academic-Self-Project

linktitle: "Academic Self and Careerism"

output:
  blogdown::html_page:
    toc: true

menu:
  docs:
    parent: "Projects"
    weight: 1
    
type: docs
lastmod: '2021-04-24T23:09:32-04:00'
featured: no

projects: []

---
# **Overview**

## **Background**

|   One part of James Daniel Lee’s argument in Which Kids Can "Become" Scientists? Effects of Gender, Self-Concepts, and Perceptions of Scientists is that girls are less likely to pursue science careers because they feel their self-concept(s) are not in line with the expectation of what a scientist is like. This suggests that, at least in part, students tend to pursue careers which they feel are congruent with their self-concept. However, this view does not account for whether college students consider their education a means to pursue their passions and express a part of their “true” self, or as a path to obtaining a lucrative career. Further, research on academic self-concept almost exclusively focuses on achievement as its main indicator or correlate. 

|   I argue that achievement is only one aspect of academic self-concept and research on this topic must account for academic gratification. That said, I seek to (1) examine whether students college major is more influenced by careerist goals or self-concepts, (2) explore how others’ perceptions (reflected appraisals) affect their self-concept and overall satisfaction with their major, (3) reveal if students who choose their major based on self-concept are more or less satisfied with their major, and (4) investigate students’ attitudes about their perceptions of their future success as it relates to the conclusions above.

## **Research Questions**

* What is the role of self-concept in students’ choice of college major? 
* What is the role of reflected appraisals in satisfaction with college major?
* Are students more likely to choose a major based on self-concept or career outcomes?
* Are students who choose a major based on self-concept more or less satisfied than those who choose a major on career outcomes?
* What is the bridge between academic self-concept and general self-concept?


## **Methods**

### Survey
Download the survey [here.](/media/Survey on Academic Self.pdf)



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


load("AcadSelfData.Rdata")
load("AcadLikertData.Rdata")


AcadSelfData$GPA <- plyr::mapvalues(AcadSelfData$GPA, from = "I don't have a GPA (First-semester freshmen)", to = "I don't know.") 
AcadSelfData$GPA <- plyr::mapvalues(AcadSelfData$GPA, from = "4", to = "4.0")

```

```{r libraries, include = FALSE}
library(tidyverse)
library(likert)
library(ggthemes)
library(extrafont)
library(plotly)
library(viridis)
```




```{r Theme setup, include = FALSE}

theme_update(text = element_text(family = "Rockwell"),
        plot.title = element_text(size = 15, face = "bold"),
        panel.background = element_rect(fill = "white"),
        panel.grid.minor.y = element_line(linetype = "dotted", color = "grey50"),
        panel.grid.major.y = element_line(color = "grey90"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor = element_blank(),
        plot.margin = unit(c(0.5, 0.5, 0.5, 0.5), units = "cm"))



threeColors <- c("#5b0909", "#34457c", "#0fb040")

fourColors <- c("#5b0909", "#571b43", "#34457c", "#0fb040")

fiveColors <- c("#5b0909", "#571b43", "#34457c", "#008a9e", "#0fb040")

sevenColors <- c("#5b0909", "#5f0720","#571b43", "#34457c", 
                 "#00558d","#008a9e","#0fb040")

eightColors <- c("#5b0909", "#5f0720","#571b43", "#4c2651", 
                 "#34457c", "#00558d","#008a9e","#0fb040")

nineColors <- c("#5b0909", "#5e0f33", "#571b43", "#4c2651", 
                  "#453568", "#34457c", "#0071a4", "#008a9e", 
                   "#0fb040")

twelveColors <- c("#5b0909", "#5f0720", "#5e0f33", "#571b43", "#4c2651", 
                  "#453568", "#34457c", "#00558d", "#0071a4", "#008a9e", 
                  "#00a07a", "#0fb040")

```




# **Sample Characteristics**

```{r Sample Characteristics}
ggplot(AcadSelfData, aes(Gender, fill = Gender)) +
  geom_bar(show.legend = FALSE) +
  geom_text(aes(label = ..count..), stat = "count", vjust = -1)+
  labs(title = "Sample Characteristics: Gender",
       y = NULL) +
  scale_fill_manual(values = threeColors) +
  ylim(0, 550)


ggplot(AcadSelfData, aes(ClassLvl, fill = ClassLvl)) +
  geom_bar(show.legend = FALSE) +
  geom_text(aes(label = ..count..), stat = "count", vjust = -1) +
  labs(title = "Sample Characteristics: Class Level",
       y = NULL,
       x = "Class Level") +
  scale_fill_manual(values = fourColors) +
  ylim(0, 225)


ggplot(AcadSelfData, aes(Majorgrouped2, fill = Majorgrouped2)) +
  geom_bar(show.legend = FALSE) +
  geom_text(aes(label = ..count..), stat = "count", hjust = -0.5) +
  labs(y = NULL,
       x = NULL,
       title = "Sample Characteristics: Area of Study",
       caption = "'Pre-prof.' indicates pre-professional fields such as pre-dental and pre-medical.") +
  scale_fill_manual(values = sevenColors) +
  ylim(0, 150) +
  coord_flip()


```





```{r}
ggplot(AcadSelfData, aes(GPA, fill = GPA)) +
  geom_bar(show.legend = FALSE) +
  geom_text(aes(label = ..count..), stat = "count", vjust = - 0.5) +
  labs(title = "Sample Characteristics: GPA",
       y = NULL) +
  ylim(0, 200) +
  scale_fill_manual(values = eightColors)



ggplot(AcadSelfData, aes(GPA2, fill = GPA2)) +
  geom_bar(show.legend = FALSE) +
  geom_text(aes(label = ..count..), stat = "count", vjust = - 0.5) +
  labs(title = "Sample Characteristics: GPA",
       y = NULL,
       x = "GPA") +
  scale_fill_manual(values = threeColors) +
  ylim(0, 425)

```




```{r Likert Scale Statements, fig.width = 8, fig.height = 8}
p <- plot(likert(AcadLikertData), col = sevenColors) +
  labs(title = "Level of Agreement Questions") +
  theme(legend.position = "none",
        legend.direction = "horizontal",
        plot.title.position = "plot",
        legend.title = element_blank(),
        legend.justification = "left",
        legend.key.width = unit(0.1, "mm"),
        legend.key.height = unit(0.35, "cm"),
        legend.key.size = unit(0.1, "mm"),
        legend.text = element_text(size = 8),
        plot.title = element_text(size = 15, hjust = 0.1),
        legend.background = element_blank(),
        panel.grid.major.y = element_line(linetype = "dotted")) +
  guides(fill = guide_legend(label.position = "bottom",
                             nrow = 1))

ggplotly(p)

```

# **Analysis**



```{r}
ggplot(AcadSelfData, aes(CareerImportant, fill = CareerImportant)) +
  geom_bar(show.legend = FALSE) +
  geom_text(aes(label = ..count..), stat = "count", vjust = -0.5) +
  labs(title = "How important was having a high-paying career\nin choosing your current major?",
       x = NULL,
       y = NULL) +
  ylim(0, 325) +
  scale_fill_viridis(discrete = TRUE,
                     direction = -1)

ggplot(AcadSelfData, aes(PassionImportant, fill = PassionImportant)) +
  geom_bar(show.legend = FALSE) +
  geom_text(aes(label = ..count..), stat = "count", vjust = -0.5) +
  labs(title = "How important was having a career you are passionate\nabout in choosing your current major?",
       x = NULL,
       y = NULL) +
  ylim(0, 450) +
  scale_fill_viridis(discrete = TRUE,
                     direction = -1)

ggplot(AcadSelfData, aes(passion_vs_career, fill = passion_vs_career)) +
  geom_bar(show.legend = FALSE) +
  geom_text(aes(label = ..count..), stat = "count", hjust = -0.5) +
  labs(title = "What is most important, going to college to get a good\ncareer or to explore your passion?",
       x = NULL,
       y = NULL) +
  coord_flip() +
  ylim(0, 300)


```



# **References**