---
title: "Experimental Survey on Hazing"
author: "C. Jackson"
date: "3/24/2021"
summary: Experimental survey on Hazing.
linktitle: "Experimental Survey on Hazing"
output:
  blogdown::html_page:
    
    toc: true
menu:
  docs:
    parent: "Projects"
    weight: 3
type: docs
tag:
 - Hazing Project
editor_options: 
  chunk_output_type: console
shiny: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = TRUE, message = TRUE, fig.show = TRUE)
```




## Data Prep and Cleaning

```{r Preparation and Cleaning, include =  FALSE}
library(tidyverse)
library(DT)
library(RColorBrewer)
library(extrafont)
library(ggpubr)
library(likert)
library(broom)
library(gmodels)
library(car)
library(DescTools)
library(qqplotr)
library(knitr)

load("HazingSurveyData.RData")
load("HazeSurveyLikert1.RData")
load("HazeScoreLikert.RData")

likertList <- colnames(Likertdf)

hazescoreList <- colnames(Likertdf3)


for(i in 1: ncol(Likertdf)){
  levels(Likertdf[,i]) <- c("Strongly disagree",
                            "Somewhat disagree",
                            "Neutral",
                            "Somewhat agree",
                            "Strongly agree")
}




HazingSurveyData$Rprt2Pol <- ordered(HazingSurveyData$Rprt2Pol, levels = c("No", "Maybe", "Yes"))




HazingSurveyData <- HazingSurveyData %>% 
  mutate(BinGlty = case_when(GltyofHazing == "Yes" ~ "Yes",
                             TRUE ~ "No/Unsure"))


###   PALETTES    ###
twelveColors <- c("#5b0909", "#5f0720", "#5e0f33", "#571b43", "#4c2651", 
                  "#453568", "#34457c", "#00558d", "#0071a4", "#008a9e", 
                  "#00a07a", "#0fb040")
fiveColors <- c("#5b0909", "#571b43", "#34457c", "#008a9e", "#0fb040")
sevenColors <- c("#5b0909", "#5f0720","#571b43", "#34457c", 
                 "#00558d","#008a9e","#0fb040")
eightColors <- c("#5b0909", "#5f0720","#571b43", "#4c2651", 
                 "#34457c", "#00558d","#008a9e","#0fb040")


```


## Theme Setup

```{r Theme, include = FALSE}
theme_update(text = element_text(family = "Rockwell", size = 12),
                         panel.background = element_rect(fill = "#f0ebd8"),
                         plot.background = element_rect(fill = "#f0ebd8", color = "black"),
                         plot.title = element_text(face = "bold", size = 15),
                         plot.title.position = "plot",
                         panel.grid.major.x = element_blank(),
                         panel.grid.minor.x = element_blank(),
                         panel.grid.major.y = element_line(color = "grey70"),
                         panel.grid.minor.y = element_line(color = "grey70", linetype = "dotted"))

```



## Sample Characteristics Plots

```{r}
ggplot(HazingSurveyData, aes(Gender, fill = BinRace)) +
  geom_bar() +
  scale_fill_manual(values = c("#350e3d", "#0fb040")) +
  labs(title = "Sample Characteristics: Gender and Race", fill = "Race", 
       y = NULL, caption = "N = 702\n- 98 respondents identified as racial or ethnic categories other than White alone.\n- 486 respondents were female.") +
  theme(legend.background = element_blank(),
        plot.caption.position = "plot",
        plot.caption = element_text(hjust = 0),
        plot.margin = margin(1, 1, 1, 1, "cm")) 




ggplot(HazingSurveyData, aes(ClassLvl, fill = fct_rev(ClassLvl))) +
  geom_bar(show.legend = FALSE) +
  scale_fill_manual(values = fiveColors) +
  coord_flip() +
  labs(title = "Sample Characteristics: Class Level", x = NULL, y = NULL) +
  geom_text(aes(label = ..count..), stat = "count", hjust = -0.8) +
  theme(plot.margin = margin(1, 1, 1, 1, "cm"),
        panel.grid.major.y = element_line(linetype = "dotted")) +
  ylim(0, 200)


ggplot(HazingSurveyData, aes(fct_rev(fct_infreq(Major)), fill = Major)) +
  geom_bar(show.legend = FALSE) +
  scale_fill_manual(values = eightColors) +
  scale_x_discrete(labels = c("No Major",
                              "Education",
                              "Other",
                              "Social Science",
                              "Humanities/Arts",
                              "Business",
                              "Biology and Health Sciences",
                              "STEM")) +
  coord_flip() +
  labs(title = "Sample Characteristics: Area of Study",
       x = NULL,
       y = NULL) +
  geom_text(aes(label = ..count..), stat = "count", hjust = -1) +
  theme(plot.margin = margin(1, 1, 1, 1, "cm"),
        panel.grid.major.y = element_line(linetype = "dotted")) +
  ylim(0, 200)
```

## Determination of Guilt and Punishment Plots

```{r Determination of Guilt}
ggplot(HazingSurveyData, aes(GltyofHazing, fill = GltyofHazing)) +
  geom_bar(show.legend = FALSE) +
  geom_text(aes(label = ..count..), stat = "count", vjust = -1) +
  ggtitle("Determination of Guilt") +
  ylim(0, 700) +
  scale_fill_manual(values = c("#350e3d", "#555a6f","#0fb040")) +
  labs(y = NULL, x = "Do you believe they are guilty of hazing?") +
  theme(plot.margin = margin(1, 1, 1, 1, "cm"))



ggplot(HazingSurveyData, aes(GltyofHazing, fill = GltyofHazing)) +
  geom_bar(show.legend = FALSE) +
  geom_text(aes(label = ..count..), stat = "count", vjust = -1) +
  ggtitle("Determination of Guilt") +
  ylim(0, 700) +
  scale_fill_manual(values = c("#350e3d", "#555a6f","#0fb040")) +
  labs(y = NULL, x = "Do you believe they are guilty of hazing?") +
  theme(plot.margin = margin(1, 1, 1, 1, "cm")) +
  facet_wrap(~ Vignette)
```


```{r}
ggplot(HazingSurveyData, aes(fct_infreq(Punishment), fill = Punishment)) +
  geom_bar(show.legend = FALSE) +
  geom_text(aes(label = ..count..), stat = "count", hjust = -1) +
  labs(x = NULL, y = NULL, title = "What Punishment should they receive?") +
  coord_flip() +
  theme(plot.margin = margin(1, 0.5, 1, 0.5, "cm")) +
  ylim(0, 250) +
  scale_fill_manual(values = sevenColors)


ggplot(HazingSurveyData, aes(fct_infreq(Punishment), fill = Punishment)) +
  geom_bar(show.legend = FALSE) +
  geom_text(aes(label = ..count..), stat = "count", hjust = -1) +
  labs(x = NULL, y = NULL, title = "What Punishment should they receive?") +
  coord_flip() +
  theme(plot.margin = margin(1, 0.5, 1, 0.5, "cm")) +
  ylim(0, 250) +
  scale_fill_manual(values = sevenColors) +
  facet_wrap(~ DrinkorPaddle)
```


## General Views about Hazing Plots

```{r Likerts, fig.height=8, fig.width=8, message=FALSE, warning=FALSE}

plot(likert(Likertdf), col = fiveColors) +
  labs(title = "Students' General Views about Hazing") +
  theme(legend.position = "top",
        legend.direction = "horizontal",
        plot.title.position = "plot",
        legend.title = element_blank(),
        legend.justification = c("left", "top"),
        legend.key.width = unit(0.5, "cm"),
        legend.key.height = unit(0.5, "cm"),
        plot.margin = margin(1, 1, 1, 0, "cm"),
        legend.text = element_text(size = 8),
        plot.title = element_text(size = 15, hjust = 0.1),
        legend.background = element_blank(),
        panel.grid.major.y = element_line(linetype = "dotted")) +
  guides(fill = guide_legend(label.position = "bottom")) 

likert.heat.plot(likert(Likertdf), low.color = "purple", high.color = "yellow") +
  labs(title = "Students' General Views about Hazing") +
  theme(plot.title.position = "plot",
        legend.title = element_blank(),
        legend.justification = c("left", "top"),
        legend.key.width = unit(1, "cm"),
        legend.key.height = unit(0.5, "cm"),
        plot.margin = margin(1, 0.5, 1, 0, "cm"),
        legend.text = element_text(size = 8),
        plot.title = element_text(size = 15, hjust = 0.1),
        legend.background = element_blank(),
        legend.position = "top",
        legend.direction = "horizontal") +
  guides(x = guide_axis(angle = 35))


```

## Creating the Hazing Score Variable

  The hazing score is a numeric representation of how positively or negatively a participant views hazing in general. The score can range from -18 to 18 and higher values indicate less positive views of hazing. The score was generated using 9 of the 15 Likert scale level of agreement items included in the survey.

* All initiations of new group members are hazing.
* An action is not hazing if the victim consents.
* Hazing is useful for group cohesion
* Hazing can be fun.
* Hazing is a serious issue which needs to be prevented.
* Hazing is dangerous
* Hazing should be illegal.
* Sometimes hazing is acceptable.
* There are no benefits to hazing new group members.


### Selecting Appriorate Variables
```{r Hazing Score Components}

Likertdf3 <- Likertdf[, -c(2,3,4,10,11,13,15)]

colnames(Likertdf3)
```

### Generating the Score

```{r}

i = 1
while(i < 10){
  
  if(i %in% c(1, 2, 3, 7)){
    
    levels(Likertdf3[,i]) <- c("2", "1", "0", "-1", "-2")
  }
  
  else{
  levels(Likertdf3[,i]) <- c("-2", "-1", "0", "1", "2")
  }
 
   i <- i + 1
}



LikertdfScore <- as.data.frame(apply(Likertdf3, 2,as.numeric))

LikertdfScore$HazeScore <- rowSums(cbind(LikertdfScore))
HazingSurveyData$Score <- LikertdfScore$HazeScore

```


### Hazing Score Plot

```{r}
ggplot(HazingSurveyData, aes(Score)) +
  geom_histogram(fill = "#571b43") +
  labs(title = "Hazing Score", 
       subtitle = "Higher scores indicate less possitive views of hazing.",
       y = NULL,
       x ="Score")
```



### Analysis

```{r}

ggline(HazingSurveyData, x = "BinGlty", y = "Score", 
    add = c("mean_se", "jitter"), 
    order = c("No/Unsure", "Yes"),
    ylab = "Hazing Score",
    xlab = "",
    title = "Do you believe they are guilty?",
    point.color = "red",
    point.size = 1,
    color = "black",
    add.params = list(alpha = 0.5))

      ggline(HazingSurveyData, x = "Gender", y = "Score", 
             add = c("mean_se", "jitter"), 
             order = c("Male", "Female"),
             ylab = "Hazing Score",
             xlab = "",
             title = "Hazing Score by Gender",
             point.color = "red",
             point.size = 1,
             color = "black",
             add.params = list(alpha = 0.5))

      
hg <- HazingSurveyData %>% 
        group_by(BinGlty) %>%
        rename("Guilty?" = "BinGlty") %>% 
        summarise(n = n(), 
                  mean = mean(Score, na.rm = TRUE), 
                  sd = sd(Score, na.rm = TRUE),
                  stderr = sd/sqrt(n),
                  LCL = mean - qt(1 - (0.05 / 2), n - 1) * stderr,
                  UCL = mean + qt(1 - (0.05 / 2), n - 1) * stderr,
                  median = median(Score, na.rm = TRUE),
                  IQR = IQR(Score, na.rm = TRUE),
                  LCLmed = MedianCI(Score, na.rm=TRUE)[2],
                  UCLmed = MedianCI(Score, na.rm=TRUE)[3],
                  `W Stat` = shapiro.test(Score)$statistic,
                  p.value = shapiro.test(Score)$p.value)

kable(hg)
   
      
hg1 <- HazingSurveyData %>%
        group_by(Gender) %>% 
        summarise(n = n(), 
                  mean = mean(Score, na.rm = TRUE), 
                  sd = sd(Score, na.rm = TRUE),
                  stderr = sd/sqrt(n),
                  LCL = mean - qt(1 - (0.05 / 2), n - 1) * stderr,
                  UCL = mean + qt(1 - (0.05 / 2), n - 1) * stderr,
                  median = median(Score, na.rm = TRUE),
                  IQR = IQR(Score, na.rm = TRUE),
                  LCLmed = MedianCI(Score, na.rm=TRUE)[2],
                  UCLmed = MedianCI(Score, na.rm=TRUE)[3],
                  `W Stat` = shapiro.test(Score)$statistic,
                  p.value = shapiro.test(Score)$p.value) 

kable(hg1)
      
  
      ggplot(HazingSurveyData, aes(sample = Score, fill = BinGlty)) +
        stat_qq_band(alpha=0.5, conf=0.95, qtype=1, bandType = "boot") +
        stat_qq_line(identity = TRUE) +
        stat_qq_point(col = "black") +
        facet_wrap(~ BinGlty, scales = "free") +
        labs(x = "Theoretical Quantiles", y = "Sample Quantiles") +
        theme(legend.position = "none") +
        scale_fill_manual(values = c("#5b0909", "#0fb040"))
      

      
      ggplot(HazingSurveyData, aes(sample = Score, fill = Gender)) +
        stat_qq_band(alpha=0.5, conf=0.95, qtype=1, bandType = "boot") +
        stat_qq_line(identity = TRUE) +
        stat_qq_point(col = "black") +
        facet_wrap(~ Gender, scales = "free") +
        labs(x = "Theoretical Quantiles", y = "Sample Quantiles") +
        theme(legend.position = "none") +
        scale_fill_manual(values = c("#5b0909", "#0fb040"))
      

  

hg2 <- tidy(wilcox.test(Score ~ BinGlty, HazingSurveyData, na.rm = TRUE, 
                       paired = FALSE, exact = FALSE, conf.int = TRUE)) 
 
kable(hg2) 
      
hg3 <- tidy(wilcox.test(Score ~ Gender, HazingSurveyData, na.rm = TRUE, 
                       paired = FALSE, exact = FALSE, conf.int = TRUE)) 
kable(hg3)      
```


